---
title: "Example plot for ED visits"
author: "Ryan Gan"
date: "4/11/2018"
output: html_document
---


Tidyverse library for ggplot and readr.

```{r}
library(tidyverse)
```

Reading in strata-specific dataframes. Jingyang, you will not need this part since you've already created the dataframes. But you will need to make sure all the variables are the same to bind the rows.

```{r}
# vector of names of files of table jingyang sent
files <- stringr::str_subset(list.files("../../data/health/"), "table")

# strata name in same order as files
strata <- c("age","total", "sex")

# reading in age results
age_results <- read_csv(paste0("../../data/health/", files[1])) %>% 
  mutate(strata = case_when(age_cat == 0 ~ "Age < 15",
                            age_cat == 1 ~ "Age 15-65",
                            age_cat == 2 ~ "Age > 65"),
         flag = ifelse(age_cat == 0 & (outcome %in% c("COPD",
                  "Cardiovascular Disease", "Ischemic Heart Disease", 
                  "Heart Failure", "Cerebrovascular Disease")), 1, 0)) %>% 
    # filter out rare events that have unstable ORs
  filter(flag == 0) %>% 
  select(outcome, strata, odds_ratio:upper95) 


# total strata
total_results <- read_csv(paste0("../../data/health/", files[2])) %>% 
  mutate(strata = "All Strata") %>% 
  select(outcome, strata, odds_ratio:upper95) 

# sex strata
sex_results <- read_csv(paste0("../../data/health/", files[3])) %>% 
  mutate(strata = case_when(sex == 0 ~ "Male",
                            sex == 1 ~ "Female")) %>% 
  select(outcome, strata, odds_ratio:upper95) 

```

Binding results dataframes together. Removing outcomes I don't want to present, and setting order of outcomes and facet wrap.

```{r}
# rowbind
results <- bind_rows(total_results, sex_results, age_results) %>% 
  # filter to certain outcomes
  filter(outcome %in% c("All Respiratory", "Asthma", "COPD", 
                        "Cardiovascular Disease", "Ischemic Heart Disease",
                        "Heart Failure", "Cerebrovascular Disease")) %>% 
  # set order of strata levels
  mutate(outcome = factor(outcome, levels = c("All Respiratory", "Asthma", "COPD", 
                        "Cardiovascular Disease", "Ischemic Heart Disease",
                        "Heart Failure", "Cerebrovascular Disease")),
         strata = factor(strata, levels = c("All Strata", "Female", "Male",
                                            "Age < 15", "Age 15-65", "Age > 65")),
         # group cvd and respiratory outcomes
         Group = as.factor(if_else(outcome %in% c("All Respiratory", "Asthma", "COPD"), 
                         "Respiratory", "Cardiovascular Disease")))
```


Small-multiples plot.

```{r}
plot <- ggplot(results, aes(x = outcome, y = odds_ratio, color = Group)) +
  geom_point() + 
  geom_errorbar(aes(ymin=lower95, ymax=upper95), width = 0.2) +
  geom_hline(yintercept = 1, linetype=2) +
  scale_color_manual(values = c("red", "blue")) +
  facet_wrap(~strata) +
  ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
  xlab("Outcome") +
   # plot theme
    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    # strip element
    strip.background = element_rect(colour=NA, fill=NA),
    panel.border = element_rect(fill = NA, color = "black"),
    # facet text size
    strip.text = element_text(size = 10),
    # axis element (tilt)
    axis.text.x = element_text(angle = 45, hjust =1),
    #axis.title.x = element_blank(),
    #axis.ticks.x = element_blank(),
    axis.title.y = element_text(angle = 90))

# print plot
plot
```

I want the plots to look like this. I may change the color scheme, but this is the first set of results. Can you please also save a high-quality image of 300 dpi and maybe 6 inches high by 8 inches wide. 
