---
title: "Counts of Asthma ED visits and SABA fills by MSA and Month"
author: "Ryan Gan"
date: "2018-05-10"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

## Introduction 

Looking and different counts and rates for asthma ED visits and SABA fills in Oregon by metropolitan statistical area (MSA) and month. I hypothesize that the Douglas-Complex fire that took place in the south-west part of Oregon impacted the MSA of Medford, 


```{r libraries}
library(tidyverse)
```

```{r read_counts}
# ed counts
ed_count <- read_csv("../../data/health/msa_month_er_counts.csv") %>% 
  rename(n_ed = n_visit) %>% 
  mutate(rate_per10k = (n_ed/POPESTIMATE2013)*10000,
         msa_name = gsub('(.*)-.*', '\\1', 
                      gsub('(.*)-.*', '\\1', gsub('(.*),.*','\\1',NAME))),
         month_name = factor(month.abb[month], 
            levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul",
                       "Aug", "Sep", "Oct", "Nov", "Dec")))
# saba counts
saba_count <- read_csv("../../data/health/msa_month_saba_counts.csv") %>% 
  rename(n_saba = n_visit) %>% 
  mutate(rate_per1000 = (n_saba/POPESTIMATE2013)*1000,
         msa_name = gsub('(.*)-.*', '\\1', 
                      gsub('(.*)-.*', '\\1', gsub('(.*),.*','\\1',NAME))),
         month_name = factor(month.abb[month], 
            levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul",
                       "Aug", "Sep", "Oct", "Nov", "Dec")))
saba_count
```

Barchart of SABA counts by MSA by month.

```{r saba_hist}
# saba histogram
saba_bar_plot <- ggplot(saba_count, aes(x = msa_name, group = msa_name,
                                        fill = msa_name)) +
  geom_bar(aes(weight = rate_per1000)) +
  scale_fill_manual("Metro Area", values = c("#3c1053", "#7f00ff", "#e100ff", 
                                             "#4389a2", "#0ed2f7", "#74ebd5")) +
  facet_wrap(~month_name) +
  ylab("SABA Fill Rate per 1,000 Persons") +
  xlab("Metropolitan Statistical Area") +
  theme_minimal() +
  theme(axis.text.x=element_text(angle = 45, hjust = 1))

saba_bar_plot
```

Alternate of month by metro area.

```{r saba_month_plot}
# saba histogram second version
saba_bar_plot2 <- ggplot(saba_count, aes(x = month_name, group = msa_name,
                                        fill = msa_name)) +
  geom_bar(aes(weight = rate_per1000)) +
  scale_fill_manual("Metro Area", values = c("#3c1053", "#7f00ff", "#e100ff", 
                                             "#4389a2", "#0ed2f7", "#74ebd5")) +
  facet_wrap(~msa_name) +
  ylab("SABA Fill Rate per 1,000 Persons") +
  xlab("Metropolitan Statistical Area") +
  theme_minimal() +
  theme(axis.text.x=element_text(angle = 45, hjust = 1))

saba_bar_plot2
```

Emergency room rates.

```{r er_month_plot}
# saba histogram second version
er_bar_plot2 <- ggplot(ed_count, aes(x = msa_name, group = msa_name,
                                        fill = msa_name)) +
  geom_bar(aes(weight = rate_per10k)) +
  scale_fill_manual("Metro Area", values = c("#3c1053", "#7f00ff", "#e100ff", 
                                             "#4389a2", "#0ed2f7", "#74ebd5")) +
  facet_wrap(~month_name) +
  ylab("Emergency Room Visits per 10,000 Persons") +
  xlab("Metropolitan Statistical Area") +
  theme_minimal() +
  theme(axis.text.x=element_text(angle = 45, hjust = 1))

er_bar_plot2
```


SABA Contrasts.

```{r saba_contrast}
saba_fit <- glm(n_saba ~ 0 + month_name + msa_name + month_name*msa_name +
                  offset(log(POPESTIMATE2013)), 
                family = poisson(), data = saba_count)

test <- MASS::glm.nb(n_saba ~ 0 + month_name + msa_name + month_name*msa_name +
                  offset(log(POPESTIMATE2013)), data = saba_count)

summary(test)
summary(saba_fit)
mean(saba_count$rate)
var(saba_count$rate)

med <- filter(saba_count, msa_name == "Medford")

saba_fit <- glm(n_saba ~ 0 + month_name +
                  offset(log(POPESTIMATE2013)), 
                family = quasipoisson(), data = med)

summary(test)
summary(saba_fit)

MASS::fitdistr(saba_count$n_saba, densfun = "negative binomial")
qcc::qcc.overdispersion.test(saba_count$n_saba, saba_count$POPESTIMATE2013, "poisson")

?MASS::glm.nb
```

